---
title: "CCLE_FINAL_20210202"
output: html_notebook
---

### Install packages 

```{r}

#install.packages("plyr")

```


### Load packages

```{r}

library(plyr)
library(tidyverse)
library(here)
library(janitor)
library(BiocManager)
library(biomaRt)
library(AnnotationDbi)
library(org.Hs.eg.db)

```

```{r}

here::here()

```

### Load data

```{r}

CCLE <- read_tsv(here::here("Data/CCLE_RNAseq_rsem_genes_tpm.txt"))

```

### Retrieve all human kinase Ensembl IDs

# Import dataset

```{r}

kinases <- read_csv(here::here("Output/kinases_update.csv"))

```

```{r}

kinases <- kinases %>% 
  dplyr::select(-X1)

```

```{r}

head(kinases)

```

# Tidy CCLE gene_id column

```{r}

CCLE <- CCLE %>% 
  mutate(gene_id_tidy = str_replace_all(gene_id, "[.][0-9]+", ""))

```

# Move tidy gene_id column to front of dataset

```{r}

CCLE <- CCLE %>% 
  dplyr::select(gene_id_tidy, everything())

```

```{r}

head(CCLE)

```

# Rename columns to match kinase_Ensembl

```{r}

CCLE <- CCLE %>% 
  dplyr::rename(gene_id = gene_id_tidy, gene_id_full = gene_id)

```

### Filter kinases in CCLE data

# Select Ensembl IDs and HGNC Names from kinase data set

```{r}

kinase_hgnc <- kinases %>% 
  dplyr::select(HGNC_Name, Ensembl_Gene_ID) %>% 
  dplyr::rename(gene_id = Ensembl_Gene_ID, hgnc_name = HGNC_Name) %>% 
  distinct(hgnc_name, gene_id)

```

```{r}

head(kinase_hgnc )

```

# Count number of kinases 

```{r}

kinase_hgnc %>% 
  distinct(hgnc_name) #619

```

# Identify duplicates

```{r}

kinase_dupes <- kinase_hgnc %>% 
	get_dupes(gene_id) %>% 
	distinct(hgnc_name, gene_id) 

```

# Join HGNC Names to CCLE data by Ensembl ID

```{r}

CCLE_hgnc <- right_join(kinase_hgnc, CCLE, by = "gene_id")

```

```{r}

head(CCLE_hgnc)

```

# Remove non-kinases (no HGNC name) from dataset

```{r}

CCLE_kinases <- CCLE_hgnc %>% 
  filter(!is.na(hgnc_name))

```

```{r}

head(CCLE_kinases)

```

```{r}

CCLE_kinases %>% 
  distinct(hgnc_name) #616

```
Three kinases appear to have been lost during the join so I will need to check this.

# List of unique kinase names from kinase_hgnc

```{r}

kin_hgnc_distinct <- kinase_hgnc %>% 
  distinct(hgnc_name) #619

```

```{r}

kin_ccle_distinct <- CCLE_kinases %>% 
  distinct(hgnc_name) #616

```

```{r}

kin_hgnc_distinct %>% 
  filter(!hgnc_name %in% kin_ccle_distinct$hgnc_name)

```

Three kinases are missing from the dataset (PRAG1, DGKK, IKBKE). I will now need to manually check to identify why these kinases are missing.

# Write .csv file 

```{r}

#write.csv(CCLE_kinases, here::here("Output/CCLE_kinases_v2.csv"))

```


### Check final kinases and check for any that are missing 

# Count number of kinases 

```{r}

CCLE_kinases %>% 
  distinct(hgnc_name) #616

```











